\documentclass[10pt]{article}
\usepackage[margin=0.85in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{array}
\usepackage{longtable}
\setlength{\parindent}{0pt}
\setlength{\parskip}{4pt}
\newcommand{\entry}[2]{\vspace{5pt}\noindent\textbf{#1}\quad #2\par}
\newcommand{\grouphead}[1]{\vspace{9pt}\noindent\rule{\linewidth}{0.4pt}\\[-2pt]{\small\textbf{#1}}\\[-4pt]\rule{\linewidth}{0.4pt}\par\vspace{2pt}}
\renewcommand{\arraystretch}{1.15}

\begin{document}
{\large\textbf{GW --- lab notebook (gravitational-wave program)}}\\
K. Bhatia \quad ORCID 0009-0007-4447-6325 \quad repo: GW (ported from kunalb541/cosmo2, E-numbering preserved)\\
Data of record: GWTC-2.1/3 PE (GWOSC/Zenodo, 76 events); GWTC-4.0 PE (Zenodo 16053484, 86 events);
LVK TGR products (Zenodo 17461225: imr/par/rin/liv); GWTC-3 cosmology icarogw (Zenodo 5645777);
PTA chains: NANOGrav 15yr fig-1 release, EPTA DR2 (Zenodo 8091568), PPTA DR3 (github).
Every battery locked-preregistered before running; every headline independently re-derived;
numbers of record in \texttt{results/}, per-battery detail in \texttt{reports/}.

\grouphead{Black-hole posterior geometry: the event atlas and its laws (E38, E40--E44)}

\entry{E38 GW posterior-geometry atlas.}{Value/shape geometry on real compact-binary posteriors:
all 76 confident GWTC-1/2.1/3 events (1.54M samples; GWOSC/Zenodo PE releases), 6 black-hole planes
each --- $(m_1,m_2)$, $(\mathcal{M},q)$, $(q,\chi_{\rm eff})$, $(M_f,a_f)$, $(d_L,\cos\iota)$,
$(z,M_{\rm tot})$; $74/76$ produce all six. The foundation for E40--E47.}

\entry{E40 GW chirp-mass lawfulness.}{Is the $(m_1,m_2)$ posterior degeneracy just the
constant-chirp-mass line? Predict each event's orientation from the constant-$\mathcal{M}$ tangent
at its median masses, vs a constant-total-mass null, 75 events. PASS: median $|\Delta\psi_{\rm
chirp}|=7.49^\circ$ vs total-mass $20.38^\circ$; chirp wins $61/75$; on the elongated subset
(axis ratio $\ge3$, $n{=}28$) $4.68^\circ$; Spearman(error, axis-ratio)$=-0.42$ (more elongated
$\to$ tighter, as predicted). A rare POSITIVE geometric-lawfulness result --- geometry is lawful
where a dominant physical combination (chirp mass) exists.}

\entry{E41 GW plane lawfulness map.}{A-priori correlation-sign predictions for all 6 planes,
derived by 12 physicist/verifier agents blind to the data, tested against measured within-event
$\rho$ (75 events). 3/6 HIT:}

\entry{E42 BH population + precision lawfulness.}{Precision test: chirp-mass fractional precision
$\sigma_{\mathcal M}/\mathcal M$ vs SNR (75 events). PASS: $r=-0.50$ ($p{=}7{\times}10^{-7}$), OLS
slope $-1.22$ (Fisher $-1$), and partial correlation with $\log\mathcal M$ at fixed SNR $=+0.94$
(the cycle-count law $N\propto\mathcal{M}^{-5/3}$). Posterior SIZE is lawful --- completing
orientation (E40) $+$ correlation (E41) $+$ size (E42). Population map: $\chi_{\rm eff}$ mean
$+0.076$, $36\%$ median-negative, $m_1$ peaks near $10$ and $35\,M_\odot$, 7 IMBH-scale remnants,
1 BNS; the geometric outliers are the physically extreme events (GW190814 $q{=}0.11$, GW191219
$q{=}0.04$ NSBH, GW200115 NSBH, GW190425 BNS).}

\entry{E43 $\chi_{\rm eff}$ per-event negatives.}{Does any event securely have $\chi_{\rm eff}<0$
(dynamical-formation signature)? From the full posteriors (75 events): $0$ events with
$P(\chi_{\rm eff}<0)>0.90$ (GW191109 sits at exactly $0.900$, 90\% CI $[-0.60,+0.12]$ spans 0),
$0$ with a 95\% CI below 0. NULL: E42's $36\%$ median-negative is largely measurement scatter, not
per-event-secure negatives.}

\entry{E44 $\chi_{\rm eff}$ hierarchical.}{Does the POPULATION require support below 0? Gaussian
fit (flat and induced-prior-reweighted, 75 events) $+$ a positive-truncation comparison. $\mu\approx
0.06$--$0.09$, $\sigma\approx0.045$--$0.061$, negative fraction median $\sim9\%$ --- but a
$\chi_{\rm eff}\ge0$ truncated population fits as well or better ($\Delta\ln L=-1.0/-0.74$), so the
negative tail is a Gaussian-tail artifact, NOT required. Reinforces E43. (Does not overturn the
LVK negative tail: a simplified pipeline --- Gaussian-only, approximate prior, no selection
function --- simply does not independently demand it.)}

\grouphead{Testing general relativity with black holes (E45--E47, E55, E57--E60)}

\entry{E45 IMR-consistency test.}{LVK GWTC-3 Tests-of-GR: the inspiral- vs post-inspiral-inferred
final mass/spin, deviation $(\Delta M_f/M_f,\Delta a_f/a_f)$, GR $=(0,0)$, 18 events. GR UPHELD ---
$D_1$ combined deviation $(-0.035,+0.002)$, GR at credible level $0.671$; $D_2$ $17/18$ consistent
(outlier GW190814 has an absent post-inspiral, a known systematic). $D_3$ (the framework's added
test): the mean-deviation directions are COHERENT (Rayleigh $p=0.001$, $13/18$ with positive
final-spin deviation), only partly along the shared degeneracy --- a coherent sub-threshold
waveform systematic that per-event tests miss.}

\entry{E46 parameterized PN deviations.}{Fractional PN inspiral-phase deviations
$\delta\hat\varphi_i$ (GR $=0$), 9 events $\times$ 10 coefficients (SEOBNRv4). GR NOT violated. The
cross-event sign-coherence test isolates \texttt{dchi0} (leading 0PN): $9/9$ events positive,
Bonferroni $p=0.004$, combined $+0.109\,[0.040,0.177]$ excluding 0 --- the documented low-PN
WAVEFORM SYSTEMATIC of a single-waveform analysis, not new physics. The non-coherent combined
exclusions (\texttt{dchi3/5l/6l}, only $2$--$4/9$ positive) are naive-product artifacts the
coherence test correctly discounts. Single waveform; multi-waveform robustness (not run) would
confirm the systematic origin.}

\entry{E47 ringdown no-hair test.}{Black-hole spectroscopy: fractional deviation of the 221
overtone frequency/damping $(\delta f,\delta\tau)$ from the Kerr prediction, GR $=(0,0)$, 22 events.
GR/Kerr UPHELD --- $D_1$ combined $0.75\sigma$ from GR ($Q=0.246$, via a robust Gaussian product);
$D_2$ $22/22$ consistent; $D_3$ ISOTROPIC (Rayleigh $p=0.79$), no coherent deviation (unlike
E45/E46 --- the coherent systematics are inspiral-specific, absent in the overtone). A first-pass
histogram-density product spuriously reported a $99.9\%$ ``no-hair violation'' that contradicted
$D_2/D_3$ (an event at $Q{=}0.987$ was only $0.25\sigma$ from GR); the Mahalanobis cross-check
caught the binning artifact and the Gaussian product corrects it.}

\entry{E55 multi-waveform robustness.}{E40's chirp-mass orientation across two independent waveform
families (IMRPhenomXPHM vs SEOBNRv4PHM, 59 events): PASS --- the law holds for both (median
$|\Delta\psi_{\rm chirp}|$ $9.1^\circ/6.1^\circ$) and the orientations agree to $3.3^\circ$
(elongated). Real geometry is waveform-robust; E46's low-PN \texttt{dchi0} coherence, by contrast,
is a known waveform systematic (would not survive). DATA-BLOCK (honest): a direct same-event
multi-waveform test of the E45/E46 GR-test coherences is impossible with GWTC-3 (the parameterized
set is FTI/SEOBNR-only; TIGER/IMRPhenom was delayed) --- would need the GWTC-2 (different events).}

\entry{E57 black-hole area theorem.}{Hawking's 2nd law from GWTC-3: $A_{\rm final}\ge A_{\rm initial}$
using the inspiral progenitors (initial areas) vs the ringdown remnant (final area), disjoint/
independent (Isi 2021 setup). Only 4 loud events carry a ringdown remnant. HELD: no violations, area
increases $5$--$79\%$, GW170104 strong single-event confirmation $0.92$, stacked $0.92\sigma$. Honest:
crude independent-pairing + broad inspiral spin priors give GW150914 $P{=}0.74$, below Isi's $0.97$ --
a consistency check, not a stringent bound.}

\entry{E58 massive graviton / Lorentz-violation.}{GW-propagation gravity test: modified dispersion
$E^2=p^2c^2+A\,p^\alpha$ over 12 events $\times$ 8 orders ($\alpha{=}0$ = graviton mass). GR UPHELD:
no Lorentz-violation/dispersion signal in any event or order. Graviton mass (per-event, robust): best
single-event $m_g < 6.7\times10^{-24}$ eV/$c^2$, median $1.8\times10^{-23}$ (LVK combined $1.27\times
10^{-23}$). Honesty catch: a naive posterior-product combined bound gave an absurd $5.9\times10^{-27}$
(3 orders too tight, grid-unstable) --- LVK use a flat-in-$A$ prior and the prior samples are empty in
the release, so it cannot be divided out; the per-event bounds are reported and the LVK combined quoted,
not a fabricated number.}

\entry{E59 ringdown mode consistency.}{Does adding the 221 overtone shift the remnant $(M_f,a_f)$ from
the 220-fundamental value? rin.zip Kerr\_220 vs plain Kerr\_221, 22 events. A naive test ``fails'' (7
events $>2\sigma$, max $6.1$), BUT the shift is COHERENT --- $M_f$ decreases in $22/22$ ($100\%$) and
$a_f$ in $21/22$ ($95\%$) with the overtone. A uniform all-event down-shift is the known ringdown-
overtone SYSTEMATIC (start-time/modeling; the GW150914 overtone controversy), NOT a no-hair violation.
The coherence lens (as in E45/E46) prevents a false violation claim; the clean no-hair result remains
E47 (221 deviation params $\sim0$, isotropic).}

\entry{E60 model-independent ringdown no-hair.}{The strongest-in-principle ringdown test: the
MEASURED 220 QNM frequency (damped-sinusoid fit, no Kerr assumption) vs the Kerr prediction
$f_{220}(M_f,a_f)$ from the INSPIRAL remnant (Berti fit) --- independent data. Kerr CONSISTENT: no
event deviates $>2\sigma$ (max $|z|{=}0.85$), combined $0.34\sigma$. GW150914 clean: Kerr $236.6$ vs
measured $246.8$ Hz ($+5\%$). Weak (only 3 events have both a DS ringdown and an inspiral remnant;
GW170814's DS frequency is broad/poorly constrained), GW150914 carries the confirmation. Ringdown
no-hair trilogy: E47 (deviation params $\sim0$) + E59 (remnant shift = systematic) + E60 (frequency
matches Kerr) all uphold the Kerr nature.}

\grouphead{Deriving the mass-plane law; the curved chirp-mass law (E65, E67)}

\entry{E65 the E40 law derived --- locked Fisher-rotation REFUTED; the posterior IS the constant-$\mathcal{M}_c$ curve.}{
Locked hypothesis: E40's residual pattern (heavy events fail toward constant-$M_\mathrm{tot}$) is a
Fisher-mixture rotation --- inspiral cycles measure $\mathcal{M}_c$, the merger band measures
$M_\mathrm{tot}$, so orientation should rotate with detector-frame mass. Model: two-direction Fisher
mixture, analytic aLIGO PSD, one constant $\kappa$ fit ONLY on the 14 locked E40-prereg events,
scored out-of-sample on the other 61. OUTCOME: D1 marginal pass ($7.45$ vs $7.72^\circ$), D2 FAIL,
D3 (parameter-free sign test) FAIL decisively --- measured $\psi$ does NOT sit between the chirp and
total tangents; rotation fractions are mostly NEGATIVE (light elongated events overshoot PAST the
chirp tangent, away from total), a sign no rotation model can produce. What the failure revealed
(post-hoc, labelled): the posterior lies along the constant-$\mathcal{M}_c$ CURVE; the principal
axis of a curved arc is its chord, not the median-point tangent. Zero-free-parameter curved law
(uses only the $\mathcal{M}_c$ median $+$ $q$ marginal): median error $4.85$ vs $7.49^\circ$
(all 75), $\mathbf{0.81}$ vs $4.68^\circ$ (elongated, $6\times$); GW190728's $14^\circ$ ``violation''
reproduced to $1^\circ$. On-record GWTC-4 out-of-sample prediction: $<2^\circ$ for elongated O4a
events. 5 contract tests; transient errno-60 read handled by retry; headlines re-derived by hand.}

\entry{E67 GWTC-4 out-of-sample: the curved chirp-mass law PASSES at $1.26^\circ$.}{The E65
prediction --- locked on record (commit 8d09cbf) from GWTC-1/2.1/3 data only, prereg pushed BEFORE
any GWTC-4 file was opened --- scored on the fully disjoint GWTC-4.0 O4a catalog (Zenodo 16053484;
86 events, all processed, zero dropped, all files h5py-validated). \textbf{D1 PASS: median
$|\psi_\mathrm{curve}-\psi_\mathrm{meas}| = 1.26^\circ$ on the 19 elongated events (threshold
$2^\circ$)} --- 18 of 19 land at $0.2$--$3.3^\circ$ while the tangent carries systematic
$4$--$11^\circ$ errors that the curvature term removes (GW231118\_090602: $10.7\to0.5$;
GW231020: $8.0\to0.2$; GW230529 mass-gap: $2.9\to0.3$). D2 PASS: curve beats tangent on all 86
($7.53$ vs $9.36^\circ$). D3: the E40 tangent law itself replicates out-of-sample ($86\%$ wins vs
$81\%$; Spearman $-0.40$ vs $-0.42$). The E65 finding is hereby PROMOTED from post-hoc to
out-of-sample confirmed: the $(m_1,m_2)$ posterior geometry of GW events is the
constant-$\mathcal{M}_c$ curve (zero free parameters --- each prediction uses only the event's own
median source $\mathcal{M}_c$ and $q$ marginal). Headlines re-derived independently + one event
recomputed end-to-end by hand (exact match).}

\grouphead{Gravitational-wave cosmology: the spectral-siren lever (E66)}

\entry{E66 spectral-siren H0 obstruction --- REFUTED at GWTC-3 precision.}{Radical claim (E61's
obstruction logic on GW): the spectral-siren H0 program is blocked by the $\sim\!33\,M_\odot$
mass-peak location $\mu_g$ acting as an uncalibrated ruler. Test: linear-response couplings on the
REAL LVK GWTC-3 joint cosmology$+$population posterior (icarogw release, $n=13{,}533$; power-law$+$
peak, restricted flat-$\Lambda$CDM). Locked stop-and-diagnose fired mid-run: the multivariate
response $C_{AB}C_{BB}^{-1}$ (partial) disagreed with the mandatory decile cross-check (marginal)
--- an apples-to-oranges comparison in the implementation, caught exactly as the two-estimator rule
intends; fixed to the marginal degeneracy track, after which all sanity gates pass. RESULT: a
$4\,M_\odot$ coherent $\mu_g$ drift moves $H_0$ by only $-0.51$ km/s/Mpc ($<1$ $\Rightarrow$
REFUTED; robust $-0.41/-0.32$ across SNR cuts) --- an order of magnitude below the CMB-vs-SH0ES gap
and inside the chain's own $\pm3.3$ statistical error. What survives: $\mu_g$ IS the single largest
systematic lever (mass group $0.38$ vs rate group $0.22$ km/s/Mpc per $1\sigma$), so the obstruction
becomes live when spectral-siren precision reaches $\sim\!1$ km/s/Mpc (O5/XG) --- then an
independently pinned or $z$-dependent peak is mandatory. Conservative caveat: the restricted
$H_0\in[65,77]$ prior compresses couplings (lower bounds), but the verdict does not hinge on it.}

\grouphead{The nanohertz band: the GW background across pulsar-timing arrays (E68)}

\entry{E68 cross-PTA GWB anatomy --- pure degeneracy-sliding; $\gamma$ marches with time span.}{
First value/shape battery in the nanohertz band, on real validated chains: NANOGrav 15yr
(HD-correlated), EPTA DR2full (HD), PPTA DR3 (CRN), + EPTA DR2new as robustness; CPTA has no
public GWB chain. D1: all pairs mutually consistent ($1.0$--$1.8\sigma$), Gaussian-faithful
(sliced-W ratios $0.98$--$1.01$). D2 HEADLINE: the constraint geometry is UNIVERSAL
($\rho(A,\gamma) = -0.85$..$-0.95$, $\psi = 110$--$116^\circ$ --- one measurement kernel) and the
pairwise mean offsets lie \textbf{94.6--99.9\%} along the common $A$--$\gamma$ degeneracy axis
(mean $0.98$; locked threshold $2/3$) $\Rightarrow$ classified degeneracy-sliding: everything the
PTAs disagree about lives along the direction a common power-law fit slides when read over
different bands. D3: the SMBHB circular prediction $\gamma = 13/3$ is excluded $>2\sigma$ by
NANOGrav ALONE ($z=+3.03$; EPTA-full $+0.29$, PPTA $+1.25$) --- the SH0ES pattern (E52): one
experiment carries the narrative; no combined number fabricated (E58 lesson). POST-HOC (labelled):
$\hat\gamma$ is STRICTLY MONOTONIC in observing span ($10.3\to24.7$ yr: $2.79\to3.25\to3.87\to4.21$),
including EPTA's internal DR2new-vs-DR2full split ($1.89\sigma$, same telescopes, offset also along
the degeneracy) --- the coherent signature of a spectrum that flattens at high frequency relative to
a single power law (or a span-correlated systematic). Successor prediction on record: longer-span
subsets of the same array should fit systematically steeper $\gamma$ (preregisterable at IPTA DR3 /
NG 20yr). 4 contract tests; every headline independently re-derived; all probe values match the
publications.}

\grouphead{GW data beyond black holes: the neutron-star EOS transfer (E63)}

\entry{E63 neutron-star mass-radius geometry.}{First transfer of the value/shape framework out of
cosmology, into the neutron-star equation of state, in the $(M,R)$ plane. Real posteriors: NICER
pulse-profile for the low-mass star J0030 (Riley 2019 X-PSI ST+PST; Miller 2019 2-oval/3-oval) and
the high-mass star J0740 (Riley 2021; Miller 2021), plus GW170817 tidal $\Lambda$ mapped to $(M,R)$
via the quasi-universal C--Love relation. Column identifications and every headline independently
re-derived; 5 contract tests pass. METHODOLOGICAL LESSON: in $(M,R)$ the axes are incommensurable
(M$_\odot$ vs km), so the degeneracy ANGLE $\psi$ is metric-dependent and meaningless (joint-whitening
forces orthogonal long axes --- a spurious $\psi$-gap$=90^\circ$ on every pair until caught); the
scale-free shape invariant is the correlation $\rho(M,R)$. RESULT D1: the NICER pipeline/model
systematic is VALUE-dominated --- every Riley-vs-Miller and spot-model difference is a central-value
offset at fixed orientation (orientation-only value fraction $0.90$--$1.00$, $\Delta\rho\le0.08$); the
sharpest is J0740, a $+1.75$~km radius offset (Miller $14.05$ vs Riley $12.30$) plus a $3.2\times$
larger Miller ellipse. This is the OPPOSITE of the cosmology analog (the KiDS statistic-dependence,
E48, was a precision/shape artifact). RESULT D3: $\rho(M,R)$ is organized by MASS REGIME and robust
across independent pipelines --- low-mass J0030 $\rho\approx+0.86$ (steep M--R banana) vs high-mass
J0740 $\rho\approx+0.25$ (flat top of the M--R curve), regime $\Delta\rho=0.61$, consistent across
pipelines ($0.84/0.89/0.86$; $0.29/0.21$) --- the NS analog of ``the kernel sets orientation''
(organizer $=$ location on the mass--radius curve). D4 (exploratory, flagged): X-ray J0030 vs
GW170817 at $\sim\!1.4\,M_\odot$ is MIXED --- both a $\sim\!1.5$~km value offset (GW prefers smaller
$R$, $\sim\!1.1\sigma$) and $\rho:0.86\!\to\!0.28$, but the GW $\rho$ is C--Love-mapping-induced. D5:
all six posteriors are near-pure ($\mathrm{Tr}\,\rho^2=0.990$--$0.996$); the $\sigma^2/8$ Chernoff
repackaging (E62) persists. Honest limits carried up front: non-Gaussian (multimodal) NICER marginals,
the mapping dependence of the GW point, and the QI repackaging.}

\grouphead{GWTC-5/O4b: the curved chirp-mass law reconfirmed out-of-sample (E71)}

\entry{E71 GWTC-5/O4b curved-law repeat.}{Third disjoint catalog after GWTC-1/2.1/3 (train) and
GWTC-4/O4a (E67). Prereg locked (commit d19a7a8) BEFORE any O4b file was opened; decision rules
byte-identical to E67, only a declared structural file/group adapter added. GWTC-5.0 O4b PE (Zenodo
20276106 $+$ 20348006): 104 per-event releases, $52.84$ GB, ZERO dropped, fully disjoint (event dates
$240413$--$250119$, zero name-overlap with the 86-event O4a set, 12 events from the 2025 tail). D1
(the on-record E65/E67 prediction; elongated axr$\ge3$, $n=32$): median $|\psi_{\rm curve}-\psi_{\rm
meas}| = 1.22^\circ < 2^\circ$ --- PASS (E67 gave $1.26^\circ$). D2: curve beats tangent, $5.41^\circ$
vs $8.23^\circ$. D3: the E40 tangent law replicates a third time (median $8.23^\circ$; chirp-vs-total
wins $88$\%; Spearman$(\mathrm{err},\mathrm{axr})=-0.69$, the strongest of the three catalogs). Zero
free parameters --- each prediction uses only the event's own median source $M_c$ and $q$ marginal.
Waveform-robust: the curve tracks the measured axis WITHIN every waveform family to $<0.7^\circ$; the
per-event largest-analysis pick (5 of 104 events do carry a Mixed group; the locked rule uses it)
shifts only the absolute orientation. Verification four legs: 17/17 contract tests; independent numpy
re-derivation of D1/D2; one event hand-recomputed from raw HDF5 via a different linear-algebra path
(match to $0.004^\circ$); computed medians reproduce the release's published PESummaryTable masses
(exact for most, to $<0.6$\% for a few late high-mass events), column identity confirmed catalog-wide.
The mass-plane geometry law is now reproducible across THREE independent catalogs spanning the full GW
history. POST-HOC (labelled): the largest curve residuals are all round posteriors (axr$\sim1$), the
ill-defined-orientation regime the axr gate excludes --- not law-breaking physics; the E72 outlier
atlas must control for axr before reading residuals as anomalies.}

\entry{E72 GW geometry outlier atlas.}{Prereg locked (0d1cebf) BEFORE any residual-axis correlation.
Where does the constant-$M_c$ curve law break? Searched the 32 elongated O4b events (axr$\ge3$;
residual $r=$ err\_curve from the locked E71 results) against 9 pre-declared axes with
Benjamini--Hochberg FDR ($q=0.05$) and partial-Spearman controlling axr; Tukey outlier fence; E59
coherence lens. RESULT --- a clean PHYSICAL NULL: none of precession ($\chi_p$), aligned spin
($\chi_{\rm eff}$), mass ratio, redshift, chirp/total mass, or secondary mass (NSBH/mass-gap proxy)
survives FDR --- the constant-$M_c$ geometry has NO source-class-dependent breakdown in O4b
(strengthens claim 1). The only surviving correlate is waveform-model disagreement (a systematic;
$\rho=-0.64$, $q_{\rm BH}=0.001$, robust to drop-one $\le0.036$), and its NEGATIVE sign says
not-physics: the events the curve fits best ($r\sim0.01$--$0.2^\circ$) are those where waveforms
disagree most --- the curve tracks the waveform-ensemble consensus better than any single waveform
tracks another (measurement optics; the E59 lesson again). One marginal Tukey outlier (GW240507,
$3.37^\circ$ vs fence $3.23^\circ$; low-SNR, asymmetric $q\sim0.27$) logged as a WEAK preregisterable
successor target, not a claim. Verify: 6 contract tests; the headline $\rho$ and the Tukey fence
independently re-derived (hand rank-Pearson; numpy percentile). Limits: $N=32$ low power; loudness axis
data-limited ($6/32$ picked groups carry a network-SNR column); eccentricity and higher-mode axes
absent from the released PE (out of scope).}

\grouphead{The loudest event: GW250114 no-hair deep-dive (E74)}

\entry{E74 GW250114 deep-dive (loudest event).}{A single-event characterization $+$ synthesis (NOT a
locked battery): the O4b TGR products are not public, so the ringdown spectroscopy and no-hair verdict
are the PUBLISHED result (Suzuki et al. 2026, arXiv:2605.03576, orthonormal QNMs; LVK IMR via GWTC-5).
GW250114 is the loudest GW event to date (network SNR $\sim80$), a near-EQUAL-mass BBH ($m_1=33.8$,
$m_2=32.2\,M_\odot$, $q\sim0.96$, low spin), remnant $M_f^{\rm src}=62.8\,M_\odot$, $a_f=0.679$ --- my
PE extraction reproduces the published IMR remnant exactly (the data gate). INDEPENDENT contribution:
propagating the remnant posterior through the Berti--Cardoso--Will Kerr fits gives the no-hair QNM
targets --- $(2,2,0)$ at $248.9$ Hz / $\tau=4.10$ ms and the $(2,2,1)$ overtone at $243.4$ Hz /
$\tau=1.35$ ms (tight to $\sim1$ Hz, the remnant being measured to $\sim1.5$\%). The published analysis
detects the $(2,2,1)$ overtone at $99.9$\% (up from $82.5$\% non-orthogonal) and finds
$\delta f_{221},\delta\gamma_{221}$ consistent with zero --- Kerr no-hair UPHELD at record SNR.
GW250114 is essentially a louder GW150914-twin: near-identical remnant ($M_f\sim68\,M_\odot$ detector,
$a_f\sim0.68$; GW150914-like $f_{220}=250$ Hz) but $\sim3\times$ the SNR, which is why the once-contested
overtone is now a clean detection. Curved-law N/A (near-equal-mass $\Rightarrow$ round posterior,
axr$\sim2$, outside the E71 elongated set). Verify: 4 QNM contract tests (GW150914 reference,
$f\propto1/M$, spin trend, overtone damps faster); published-value validation of all four remnant/mass
quantities. Limit: the strain spectroscopy is not reproduced --- the no-hair verdict is the cited
published result; the area theorem awaits the O4b TGR release.}

\grouphead{A geometric test of GR: the inspiral mass-combination exponent (E78)}

\entry{E78 posterior-geometry test of GR (the inspiral exponent).}{EXPLORATORY/post-hoc, grown from
E71/E72: since the $(m_1,m_2)$ posterior orientation is a zero-parameter function of the constant-$\mathcal{M}_c$
curve, INVERT the law and measure which chirp-mass exponent the geometry prefers --- a test of GR through
posterior geometry, not waveform templates. Generalize the curve to $\mathcal{M}_p=(m_1 m_2)^p(m_1+m_2)^{1-2p}$
(GR: $p=3/5=0.600$, set by the 0PN phase $\propto\mathcal{M}_c^{-5/3}$); the orientation is SCALE-FREE
($m_1(q;p)\propto q^{-p}(1+q)^{2p-1}$), isolating $p$ from the mass scale. On the 32 elongated O4b events:
$\hat p = 0.616\pm0.011\,(\mathrm{stat})\pm0.004\,(\mathrm{waveform})$ vs GR $0.600$ --- CONSISTENT at
$1.3\sigma$, a $\sim2$\% geometry-only confirmation of the inspiral mass-combination exponent. Validated:
injection recovers the true exponent unbiased (bias $+0.0002$; $0.55\to0.550$, $0.65\to0.651$); the waveform
systematic is negligible ($\pm0.004$ --- the per-event $\sim3^\circ$ scatter is incoherent and averages out);
5 contract tests; sensitivity $\mathrm{d}\psi/\mathrm{d}p\sim30^\circ$/unit. A deviation would have signalled
modified inspiral phasing (dipole radiation / varying-$G$); none seen. HONEST: a methodologically independent
cross-check, NOT more sensitive than the LVK $-1$PN dipole bound; O4b-only; the $\sim1.2^\circ$ E71 law
residual sets a per-event floor ($\sim0.04$ in $p$), so the small $+0.016$ offset is at the level of the law's
own imperfection, not claimed as physics. Successor (preregisterable): the fitted exponent is reproducible
across GWTC-3/O4a/O4b and $=0.600$.}

\entry{E79 cross-catalog exponent --- a $3\sigma$ ``deviation'' that is a systematic.}{Tests E78's successor
(prereg locked BEFORE O4a was opened): is the geometry-fit exponent reproducible across catalogs and
$=0.600$? On GWTC-4/O4a (86 events, disjoint from O4b): $\hat p(\mathrm{O4a})=0.647\pm0.014$ ($3.3\sigma$
from GR), $\hat p(\mathrm{O4b})=0.616\pm0.011$; they REPRODUCE each other (D2 PASS) but combined
$\hat p=0.628\pm0.009$ sits a naive $3.1\sigma$ above $0.600$ (D1 FAIL). The prereg pre-committed a
follow-up before any claim --- and it dissolves the anomaly: (i) the two independent catalogs disagree by
$0.031\approx$ the offset $0.028$, so the bootstrap stat error underestimates --- a systematic floor
$\pm0.016$ gives $1.5\sigma$; (ii) the offset is elongation-dependent (Spearman$(p^*,\mathrm{axr})=-0.41$,
coherent in BOTH catalogs) and the cleanest, most-elongated half gives $\hat p=0.606\approx$ GR. By the
coherence lens a cross-dataset-coherent anomaly is a SYSTEMATIC --- here a finite-width / higher-order
imperfection of the pure-0PN constant-$\mathcal{M}_c$ model, vanishing in the clean limit --- NOT a GR
violation. Honest headline: the effective exponent is $0.628\pm0.009\,(\mathrm{stat})\pm0.016\,(\mathrm{syst})$,
GR-consistent at $1.5\sigma$, cleanest events $0.606$; the naive $3\sigma$ is a false alarm the discipline
caught (the E47/E58 lesson, live). Verified: injection unbiased on both catalogs; disjointness $=0$;
column-ID clean; honest budget independently re-derived. Successor: a 1PN geometric model should reproduce
the $-0.4$ elongation trend and flatten it to $0.6$; add GWTC-3 as a third point.}

\grouphead{The information anatomy of a detection (E73)}

\entry{E73 information anatomy of a detection.}{CHARACTERIZATION (forward Fisher from the waveform model
$+$ PSD, NOT strain-extracted; the data-driven test is the successor). For all 190 real O4a$+$O4b events,
the frequency by which $50\%$/$90\%$ of the Fisher information about each parameter is accumulated ---
chirp mass ($0$PN, $\sim f^{-5/3}$), mass ratio ($1$PN), spin ($1.5$PN). RESULT: the ordering
$\mathcal{M}_c<\eta<\chi$ holds for ALL 190 events --- chirp mass is always learned first ($\sim35$ Hz),
spin last; the anatomy ``richness'' ($\chi_{f50}-\mathcal{M}_{c,f50}$) has Spearman $=-1.00$ with total mass
(lighter / longer-inspiral $\Rightarrow$ wider anatomy; heavy events merge almost immediately in band and
compress it). A frequency-resolved ``when is each thing measured'' anatomy of every detection. GR-test
direction (a $-1$PN dipole term): the RANK is reliable (Spearman $+1.00$ with mass; the lightest events ---
GW230529, the mass-gap merger, and GW230518 --- are most sensitive, matching the known low-mass preference
of dipole tests), but the absolute forecast is conditioning-limited (Fisher cond $\sim10^{12}$; arbitrary
$\beta$ units), so only the ordering is quoted. Verify: 5 contract tests ($\mathcal{M}_c<\eta<\chi$
ordering, $f\propto1/M$, $\sigma_\beta\propto1/\rho$, lighter$\to$tighter dipole); conditioning explicitly
checked; ordering $+$ richness independently re-derived; suite $37/37$. Honest ceiling: model-PREDICTED, not
strain-MEASURED --- the data-driven test (re-analyse loud events in frequency slices, compare empirical vs
GR-Fisher information accumulation) is the successor, along with a numerically-stable ppE dipole forecast
versus the LVK $-1$PN bound.}

\grouphead{The GW inference manifold: the catalog as one geometric object (E82)}

\entry{E82 the GW inference manifold.}{CHARACTERIZATION (post-hoc; radical idea made concrete). Treat all
190 O4a$+$O4b $(m_1,m_2)$ posteriors as ONE object --- each a Gaussian $N(\mu,\Sigma)$, the catalog the
cloud of them, distance $=$ 2-Wasserstein (location $|\Delta\mu|^2$ $+$ Bures shape). RESULT: (1)
location(population) 95\% vs shape(measurement) 5\% --- mass-range-dominated, partly trivial; (2) the
curved law slaves each ellipse's orientation to its location (median $1.26^\circ$, E71; residual
physics-blind, E72), so orientation is NOT an independent axis; (3) classical-MDS variance 97\%/3\%/0\%,
participation-ratio effective dimensionality $=1.07$ (naive per-event DOF $=5$); axis 1 $=$ POPULATION
($|\rho_{\mathrm{Mc}}|=0.99$), thin axis 2 $=$ PRECISION ($|\rho_{\mathrm{axr}}|=0.51$). VERDICT: the
catalog of black-hole mergers, geometrically, is a population sequence (chirp mass) dressed with a thin,
law-determined measurement structure --- subtract the law and the population remains; measurement and
population are geometrically separable, and here separated. The catalog-level form of the program's thesis
(E71 orientation $=$ law(location); E72 residual physics-blind): the measurement optics occupy, and
REMOVE, exactly one geometric axis. Verify: 4 contract tests (Bures closed form vs scipy \texttt{sqrtm} to
$10^{-8}$; MDS recovers synthetic 1D/2D dimensionality); suite 41/41. Honest: the 95/5 split and
eff-dim$\sim$1 are mass-range-dominated (not oversold); Gaussian approximation; mass plane only. Successor:
the full-parameter-space (spin, distance) Fisher-Rao version --- the deferred E80 sloppy-spectrum program.}

\grouphead{A strain-level pipeline, validated on GW250114 (E83)}

\entry{E83 from-scratch strain pipeline.}{CAPABILITY $+$ consistency check (``science first''): the
strain-level analysis the exotic-test class requires, built with numpy/scipy only (no gwpy). Download
GWOSC O4b strain $\to$ whiten (off-source Welch PSD) $\to$ bandpass $\to$ instantaneous frequency (Hilbert)
for the chirp; post-peak dominant-frequency $+$ log-envelope damping for the ringdown. GW250114 (H1):
whitened peak $388\sigma$ (loudest to date); the CHIRP is recovered ($f_{\mathrm{inst}}$ rising through
merger $89\to117\to187\to220$ Hz across $\pm5$ ms) and the RINGDOWN is recovered $f_{\mathrm{dom}}=241$ Hz
($3.2$\% from E74's Kerr $248.9$ Hz), $\tau=4.7$ ms (E74 $4.10$) --- CONSISTENT with the E74 Kerr
prediction. This independently cross-checks E74: the strain ringdown matches the Kerr QNM computed from the
INSPIRAL-derived remnant --- inspiral$\to$remnant$\to$ringdown closes in the actual data. Verify: 4 contract
tests (a $250$ Hz$/4$ ms damped sinusoid injected into noise is recovered; whitening flattens red noise;
the bandpass suppresses out-of-band tones $>1000\times$; a chirp's instantaneous frequency rises); suite
45/45. HONEST: a consistency check, NOT a measurement of record --- single-detector, single-mode,
fixed-start point estimate (few-\% frequency / $\sim$15\% $\tau$ bias vs a proper Bayesian multi-mode
ringdown). Successor: a Bayesian ringdown; the data-driven frequency-geometry test (E73 successor); the
exotic echo / dark-matter-frequency-rotation tests ONLY after multi-event validation $+$
coherence-lens/systematic controls (per E79/E72, a positive there is likelier a systematic than new
physics).}

\entry{E84 data-driven chirp mass --- ballpark result, honest bias, feasibility boundary.}{First physics
from the E83 pipeline: fit the textbook 0PN law $f^{-8/3}(t)=K(t_c-t)$ to the spectrogram RIDGE of
GW250114's whitened H1 strain --- a template-free chirp mass. Raw: $\mathcal{M}_c^{\rm det}=33.0\,M_\odot$
vs PE $31.1$ ($+6$\%). BUT injection calibration shows the estimator is biased $+15$--$30$\% HIGH
(frequency quantization $+$ window smearing in an accelerating sweep; grows with $\mathcal{M}_c$), so the
raw agreement is PARTLY FORTUITOUS --- honest claim: ballpark-consistent with a $\sim$20--30\%
template-free systematic, NOT precision. (The double-check caught this before it was banked --- the E79
lesson again; a scipy STFT window-center double-offset bug was also caught by the contract tests.)
NEGATIVE RESULT with mechanism: the band-resolved version (the frequency-geometry observable) is
INFEASIBLE via ridge methods --- GW250114 sweeps $64\to200$ Hz in $<25$ ms, less than one STFT window
(upper bands: 4 and 1 confident points), while light events are too faint per-slice; this is E73's
anatomy COMPRESSION observed in data. The data-driven frequency-geometry test therefore requires
band-limited Bayesian PE (a major build), not more signal processing. Verify: 5 contract tests
(slope-inversion roundtrip; recovery within the characterized bias envelope, never low; bias GROWS with
$\mathcal{M}_c$; noise false-positive rate $<15$\%); suite 50/50; the injection-bias table ships in the
results JSON.}

\entry{E85 Bayesian ringdown of GW250114 --- \textbf{RETRACTED, superseded by E87.}}{\textbf{The numbers in
this entry are withdrawn: the posterior was PRIOR-DOMINATED (the whole $160$--$340$ Hz grid sat within
4 lnL, so the ``$90$\% CI'' was the grid) and was statistically indistinguishable from no signal. Two
covariance defects cost ${\sim}33\times$ in matched-filter SNR; see E87 for the diagnosis, the fix, and
the validated measurement. The machinery below --- analytic amplitude/phase marginalization, the Kerr
inversion, and the truncation that defeats acausal whitening leakage --- is sound and is reused unchanged.
Retained verbatim as the record of the error.} The
Bayesian layer over the E83 pipeline: posterior on $(f,\tau)$ for the 220 mode $\to$ BCW inversion $\to$ a
ringdown-ONLY $(M_f,a_f)$ posterior vs the IMR remnant. Final likelihood: TRUNCATED time-domain
(pyRing-style) --- the window is compared to the model under the empirical noise AUTOCOVARIANCE (Toeplitz
$+$ Cholesky), amplitude/phase marginalized analytically (linear params), $(f,\tau)$ gridded exactly
($\sim$1 s/fit, deterministic). RESULT (H1-only, 220-only): $f_{220}=254\,[167,333]$ Hz (start$+3$ ms) and
$256\,[169,333]$ Hz (start$+5$ ms) --- start-time STABLE, median at the IMR-Kerr prediction ($248.6$ Hz);
ringdown-only remnant $M_f=79\,[45,122]\,M_\odot$, $a_f=0.87\,[0.19,0.98]$ --- the IMR remnant
$(68.4,0.679)$ lies INSIDE the 90\% region. A no-hair consistency check closed end-to-end with our own
machinery, at the honest precision of a single-detector single-mode segment (ringdown SNR $\sim6$). The
build defeated three diagnosed artifacts, each now regression-guarded: (1) an amplitude prior that
EXCLUDED the true $\sim2\times10^{-21}$ strain scale; (2) an ensemble-sampler convergence failure (fixed
by analytic linear-amplitude marginalization $+$ exact gridding); (3) the important one ---
whiten-then-window ACAUSAL LEAKAGE: the $388\sigma$ merger rings forward through the PSD structure into
the window, manufacturing start-time-unstable fake modes (192 vs 273 Hz); cured by the truncated
autocovariance likelihood, and guarded by a regression test in which a $500\times$-noise pre-window burst
must yield a flat likelihood surface. Verify: 6 contract tests (sampler on a known Gaussian; BCW roundtrip
to $10^{-6}$; grid injection recovery; surface peaks at injection; leakage regression); suite 56/56.
Limits: H1-only, 220-only, fixed start, weak-consistency precision. Successor: L1 $+$ coherent network,
221 overtone (4 linear amplitudes, same marginalization), start-time marginalization --- then
overtone/area-theorem questions in-house, and only then the exotic-echo class on this artifact-safe
foundation.}

\entry{E86 the band-resolved boundary, closed from both sides.}{E84 found band-resolved $\mathcal{M}_c$
infeasible on a HEAVY event and PREDICTED the complementary light-event failure; E86 tests that on the best
case in the catalog and closes the method. Event: GW241011 --- the MOST ELONGATED event (axr $17.3$,
curved-law residual $0.62^\circ$), $q=0.32$, SNR$\sim$36, and a $\sim$1.2 s inspiral ($7\times$
GW250114's), H1$+$V1 public. RESULT: it fails here too. (1) Only the LOWEST band tracks a real chirp ---
ridge monotonicity Spearman$(t,f)$ is $+0.82$ at 30--50 Hz but $-0.34$, $-0.02$, $-0.38$ at 40--70, 50--90,
80--150 Hz: line/noise structure, not a chirp (the conditioning already whitens AND notches a measured line
forest --- 60.00 Hz sits $\sim$400$\times$ above the local median). That is E84's predicted
per-slice-SNR failure. (2) The natural rescue --- a LONG (0.5 s) window, affordable only because the
low-frequency sweep is slow --- buys monotonicity but is DEGENERATE: injections of known chirps into REAL
off-source noise give recovered $7.60$, $7.89$, $4.88$ for true $8.0$, $10.9$, $14.0$, i.e. a $36$\% change
in truth moves the estimate $4$\% --- \textbf{response slope $0.11$} ($1.0$ unbiased, $0$ no information).
The real $\mathcal{M}_c^{\rm det}=8.35$ is ``consistent'' with the true $\sim$10.9 only vacuously. VERDICT:
ridge/spectrogram methods cannot deliver band-resolved $\mathcal{M}_c$ for ANY event class --- heavy (sweep
$>$ window), light (faint per-slice $+$ lines), long-window rescue (degenerate) --- so the
frequency-geometry observable REQUIRES matched-filter or banded Bayesian PE, now a measured requirement
rather than a guess. Verify: 5 contract tests (notch suppresses a planted 60 Hz line; monotonicity
discriminates injected chirp from noise; estimator $\sim$unbiased where the window IS adequate; and the
compressed response slope is itself regression-guarded at $<0.75$); suite 61/61.}

\entry{E87 E85 retracted --- its ringdown posterior WAS the prior; covariance fixed, validated vs LVK
pyRing.}{E85 reported $f_{220}=253.9\,[167,333]$ Hz as an ``honestly wide'' Kerr-consistent measurement. It
was not a measurement: the entire grid sat within \textbf{4 lnL}, so the interval was the grid, the IMR
remnant fell inside because EVERYTHING did, and the ``start-time stability'' was a flat surface. THE TELL:
adding L1 to H1 failed to tighten the posterior ($162.0$ vs $160.0$ Hz) --- impossible if there were
information to add. Diagnosis: the fraction of the $(f,\tau)$ grid within 2 lnL of the peak was
\textbf{1.00} on real data (0.02 on a loud injection), and both detectors railed at the grid's LOW edge.
ROOT CAUSE, two defects in the truncated-likelihood noise covariance, together ${\sim}33\times$ of lost
matched-filter SNR: (1) an ACF taper applied to force positive-definiteness crushed the off-diagonal
correlations while keeping $\mathrm{acf}[0]$, driving $C$ toward white noise at the BROADBAND rms of
high-passed strain, so a 250 Hz filter was charged 25--60 Hz seismic power ($8\times$); (2) a diagonal
ridge of $10^{-3}\times$ the BROADBAND variance was itself a white floor far above the true in-band noise
(a further ${\sim}3.5\times$). FIX: build the ACF as the Wiener--Khinchin transform of a Welch PSD ---
positive-definite BY CONSTRUCTION, so no taper is needed and the noise colour survives --- with the ridge
at $10^{-6}$. A bug introduced WHILE fixing it and caught by an explicit check: the one-sided PSD must be
halved before the inverse FFT, or $C$ comes out exactly $2\times$ too large, inflating every credible
interval by $\sqrt{2}$ WITHOUT moving the peak (invisible to any peak-location test; caught only by
testing $\mathrm{acf}[0]$ against $\mathrm{var}$(noise): ratio $2.0012\to1.0006$). VALIDATION GATE, the
point of the exercise, PASSED: against LVK's own pyRing Kerr\_220 posteriors at 19 start times (Zenodo
17018009), mapped to $f_{220}$ through the SAME BCW relation --- median offset \textbf{2.0 Hz} over
2--18$M$ (max 5.8), LVK's median inside our 90\% CI at \textbf{10/11} start times, and the credible-interval
WIDTHS track ($16.4$ vs $16.3$ Hz at $12M$), i.e. the covariance is CALIBRATED, not merely centred.
PHYSICS: the network start-time scan reproduces the early-time bias of a single-220 fit --- $230$ Hz with
an \textbf{8 Hz} CI at $t_{\rm peak}$ (tight and WRONG), rising to a $251$--$254$ Hz plateau at 12--18$M_f$,
then back to the prior past $24M_f$. \emph{The tightest interval is the one that most confidently excludes
the right answer.} Also withdrawn: a ``$251.7^{+5.1}_{-5.0}$ Hz LVK figure'' used earlier in the session was
never sourced and is not in the release; all comparison now reads LVK's actual samples at runtime. Verify:
9 contract tests (WK normalization; PD-without-taper; corrected covariance $>5\times$ the E85 construction;
pure noise gives an UNINFORMATIVE posterior; inherited pre-window-burst leak guard; injection recovery;
adding a detector must not broaden; Kerr roundtrip; peak location) --- with the synthetic-noise fixture
CALIBRATED (a gentle $60\times$ spectral step reproduces only a $1.8\times$ taper penalty vs ${\sim}20\times$
for real H1, so it was steepened until it reproduced ${\sim}40\times$; a fixture too tame to exhibit the bug
would make the regression test vacuous --- the same failure mode as the bug). Suite 70/70.}

\entry{E88 the 221 overtone in GW250114 --- a significant two-mode preference that is NOT the overtone.}{
A negative with a named cause. GEOMETRY FIRST, computed before looking at any published result: at the IMR
remnant the 221 sits $5.5$ Hz from the 220 ($243.1$ vs $248.6$) but decays $3\times$ faster ($1.35$ vs
$4.10$ ms), so over a 40 ms window the bases are \textbf{${\sim}76$\% collinear} --- most of the overtone is
something the 220 can already absorb, and only the ORTHOGONALIZED SNR$_\perp$ is available:
$6.9$ at $t_{\rm peak}$, $3.3$ at $3M$, $0.9$ at $8M$. So any overtone evidence must live at 0--3$M$ and be
gone by $6M$; this independently predicts the profile of LVK's own evidences. TWO METHOD RESULTS were
needed first: (a) a Bayes factor needs a PROPER amplitude prior --- E85/E87 marginalized amplitudes with a
flat, improper one (fine for a posterior, undefined for model comparison); replaced by a Gaussian prior,
formula verified against brute-force 2-D integration to $3\times10^{-14}$. (b) The naive multi-detector
model DESTROYS the test: a free 221 amplitude per detector adds 4 linear parameters and costs
\textbf{${\sim}11$ lnB of pure Occam penalty on signal-free data}. But the 221 and 220 share a detector's
antenna pattern, so the complex ratio $\rho=A_{221}/A_{220}$ is COMMON across detectors; writing
$\mathrm{Re}[A_{\rm det}(T_{220}+\rho T_{221})]$ keeps it linear in $A_{\rm det}$ via
$M_{\rm red}=M_{\rm full}K(\rho)$ (exact to $8\times10^{-16}$). Both models then carry two linear
amplitudes per detector: $\sigma_A$ cancels (at most $0.32$, median $0.13$, over a $33\times$ range), the
whole overtone cost is the $\rho$ prior volume (synthetic injections with $\rho$ contained:
$2.00/2.12/1.83$ vs $\ln 9=2.20$; real data $1.01$--$2.06$, median $1.52$, i.e.\ $\rho$ is only weakly
measured), and the signal-free penalty drops to $+0.9$.
DOMINANT SYSTEMATIC, measured: with the remnant PINNED an injected $|\rho|=1$ overtone gives
$\Delta\ln B=+21.2$; with $(M_f,a_f)$ MARGINALIZED it collapses to $-0.05$ --- \textbf{essentially ALL
absorbed} (at $|\rho|=2$: $+110.0\to+15.1$, 86\%). The
single-mode model just moves the remnant, peaking at $(54.0,0.150)$ whose 220 mode is $f=234.8$ Hz,
$\tau=3.04$ ms. \emph{That is E87's early-time bias seen from the other side: same mechanism, two
batteries.} Window length does not rescue it ($\Delta\ln B\sim0$--$1$ for 20--320 ms). RESULT:
$\Delta\ln B = 6.15/6.09/6.33/5.88$ at $0/1/2/3M$, still $4.94$ at $8M$, negative by $10M$. Against the
CORRECT null --- a loud 220 with NO overtone injected into real noise, 13 realizations, $-2.06\pm0.98$ ---
the observed $+6.15$ is \textbf{8.4$\sigma$}, 0/13. (The signal-free null, $+0.91\pm0.58$, is the WRONG one:
it contains no 220 either and would flatter the result to $9.4\sigma$.) THE DISCRIMINATOR: a genuine 221
must fade with a 2.01 ms e-folding, and an injected $|\rho_0|=1.5$ template duly falls $+5.69$ from $0M$ to
$8M$ (undetectable already by $1M$); the observed preference falls only $\mathbf{+1.21}$, persisting to $8M$
where SNR$_\perp=0.94$ and no overtone can survive. \textbf{A preference that does not decay like an
overtone is not the overtone} --- by the standing coherence lens, a systematic of roughly constant size
${\sim}+5$. LVK's profile DOES decay ($20.8\to10.1\to5.7\to0.7$ over 0--3$M$), matching the SNR$_\perp$
prediction; theirs, not ours, is the analysis that supports an overtone claim. We match them at 2, 8, 10
and 12$M$ and differ most at 0--1$M$, where our looser detector model costs us. The systematic is NOT
identified; sky-position-tied amplitudes are the first suspect. Verify: 10 contract tests; suite 80/80.}

\entry{E89 spin magnitude rises with mass across 265 events --- a hierarchical-assembly signature.}{
THE PHYSICS: a comparable-mass BBH merger leaves a remnant with $a_f\approx0.69$ almost independently of
what went in --- one of GR's cleanest predictions, and visible directly in this catalog (\texttt{final\_spin}
sits at $0.690$--$0.707$ in EVERY mass bin). Black holes that are themselves merger remnants should
therefore carry $a\sim0.7$ and sit preferentially at high mass. WHY E42--E44 SAW NOTHING: those batteries
hunted this channel in $\chi_{\rm eff}$ and found no requirement for it, which is exactly right ---
dynamical assembly gives ISOTROPIC orientations, so large spin magnitudes average to $\chi_{\rm eff}\approx0$.
$\chi_{\rm eff}$ is structurally blind to the population; the magnitude $a_1$ is not. This is E44's question
asked in the observable that can see it, on 265 events instead of 75. RESULT: hierarchical fit of
$p(a_1|m_1)=\mathrm{Beta}(\mathrm{mean}=\mathrm{sigmoid}(b_0+b_1\ln(m_1/30)))$ gives $b_1=\mathbf{+1.193}$,
$\Delta\ln L=10.94$, population mean $a_1$ running $0.083\to0.304\to0.499\to\mathbf{0.647}$ at
$m_1=8/30/60/100\,M_\odot$ against the $0.69$ remnant reference. Descriptive Spearman $+0.554$
($p=9\times10^{-23}$), repeating independently in GWTC-3 ($+0.683$), O4a ($+0.490$) and O4b ($+0.475$). The
$a_1$ prior is uniform$(0,1)$ --- verified against the \texttt{priors} group --- so each marginal posterior
is proportional to its marginal likelihood. SIGNIFICANCE against nulls, not $\chi^2$, because heavier events
are genuinely less informative (Spearman$(\ln m_1,$ concentration$)=-0.517$): forward simulation with
mass-independent truth gives $b_1=-0.046\pm0.138$; permuting masses within concentration strata gives
$-0.002\pm0.173$. Both sit at ZERO, so the estimator is not biased by the confound, and the observed
$+1.193$ exceeds the maximum of all \textbf{300} null trials. REMNANT-SPIN LEAK REFUTED: a leak would put
heavy $a_1$ at \texttt{final\_spin} (gap ${\sim}0$, per-event correlation ${\sim}+1$); the measured gap stays
$-0.197$ above $60\,M_\odot$ and the per-event correlation is ${\sim}0.4$--$0.5$ for LIGHT events too.
TWO TRAPS, both now contract-tested: (1) an ``informative events'' cut is CIRCULAR --- among heavy events
spin is well measured precisely WHEN IT IS LARGE (Spearman$(a_1,$KS-from-prior$)=+0.93$ across the 10
heaviest), and the cut manufactures the very trend it appears to confirm; not used in the result. (2) The
first null generator was WRONG and would have made this battery UNDER-report itself: modelling a posterior
as a Beta centred on the truth makes a small true spin pile against $a_1=0$ and look informative, whereas
real uninformative posteriors relax to the FLAT prior (real heavy low-information events sit at median
$a_1=0.474$). The wrong shape produced a spurious null slope of $+0.42\pm0.11$ that would have been
subtracted from $+1.19$ as estimator bias; fixed to a truth/uniform mixture. A tempting analytic argument
--- that a flat posterior contributes $\ln 1=0$ to every model and so cannot bias anything --- holds only
in the exact-flat limit; NEAR-flat posteriors do not obey it, which is why the nulls had to be run rather
than reasoned about. LIMIT, and it is not small: \textbf{SELECTION EFFECTS ARE NOT MODELLED} --- this is the
DETECTED population, not the astrophysical one, and aligned spin raises SNR, so the astrophysical amplitude
is not established without a VT/injection campaign. The effect is also not novel; what is new here is the
independent measurement with the artifact and null calibration spelled out. Verify: 9 contract tests;
suite 89/89. Successor: a 1G$+$2G mixture with a selection function, and the tilt-isotropy test.}

\entry{E90 superradiance / ultralight bosons (SCOUT, not committed) --- no exclusion survives calibration.}{
Penrose's rotational-energy extraction in its wave form: an ultralight boson forms a bound cloud around a
spinning BH and spins it down, so if a boson of mass $\mu$ exists every BH must satisfy
$a<a_{\rm crit}=4\alpha/(1+4\alpha^2)$ with $\alpha=GM\mu/\hbar c^3$ (derived from scratch; $\alpha=0.7485$
for $M=10\,M_\odot$, $\mu=10^{-11}$ eV vs the literature's ${\sim}0.75$). Applied to the 265 E89 posteriors,
a naive analysis ``excludes'' $\mu\in[2\times10^{-14},4\times10^{-12}]$ eV. FOUR successive controls each
killed a version: (i) $P(a>a_{\rm crit})$ is prior-dominated (the uniform spin prior alone gives
$0.5$--$0.95$); (ii) a Bayes factor against a UNIFORM null makes the boson look FAVOURED ($\ln B=+40$), but
shuffling spins across events also gives $+16$ --- real spins are simply low, so ANY truncation wins; (iii)
using the E89 astrophysical population as the null (the right structure) gives $\ln B=-1403$, but
$-690=\ln 10^{-300}$ is the numerical floor: single events' posterior TAILS, exactly where PE is least
trustworthy; (iv) with an E79-style systematics floor ($\epsilon=0.1$, capping any one event at $\ln 0.1$),
$\ln B=-69$ --- but mock catalogs drawn from the E89 population with \emph{no boson at all} give
$-63.0\pm2.8$. VERDICT: the real data sit $\mathbf{1.95\sigma}$ below the no-boson expectation; the
``excluded band'' was a look-elsewhere artifact across 60 $\mu$ values. \textbf{No boson mass is excluded.}
Kept as a scout (no src/results committed): the reusable content is the $\alpha$/$a_{\rm crit}$/$\tau_{\rm SR}$
machinery and the demonstration that E89's population is the correct null for exotic-spin tests.}

\end{document}
