\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{natbib}
\setcitestyle{numbers,square}

\title{\textbf{The Geometry of Gravitational-Wave Inference:\\
a lawful anatomy of posteriors, systematics, and a geometric test of general relativity}}
\author{Kunal Bhatia\\
\small Independent researcher, Meerut, India \\
\small ORCID: 0009-0007-4447-6325}
\date{Draft --- \today}

\begin{document}
\maketitle

\begin{abstract}
The parameter-estimation posteriors of compact-binary gravitational-wave events have a shape that is
usually treated as a nuisance. We show instead that this shape is \emph{lawful}: the principal-axis
orientation of each source-frame $(m_1,m_2)$ posterior is quantitatively predicted, with \emph{no fitted
parameters}, by the constant-chirp-mass curve evaluated at the event's own median chirp mass over its
mass-ratio marginal. The local constant-$\mathcal{M}_c$ tangent --- the Fisher/metric approximation
underlying rapid parameter-estimation tools --- leaves residuals that grow with posterior elongation; the
curve (chord) term removes them. Derived on GWTC-3 and locked before opening later catalogs, the law
predicts the measured orientation out-of-sample on the disjoint GWTC-4.0/O4a and GWTC-5.0/O4b releases to a
median mismatch of $1.26^\circ$ and $1.22^\circ$ respectively. A physics-blind search finds \emph{no}
source-class dependence of the residuals: precession, spin, mass ratio, redshift, mass and neutron-star--black-hole
character all fail to correlate after multiple-comparison control; the only surviving correlate is
waveform-model disagreement, a systematic. Inverting the law to \emph{measure} the leading inspiral
mass-combination exponent yields $0.628\pm0.009\,(\mathrm{stat})\pm0.016\,(\mathrm{syst})$, consistent with
the general-relativistic value $3/5$; a naive analysis produced a spurious $3\sigma$ ``deviation'' that a
coherence and systematic-floor analysis correctly demotes to $1.5\sigma$. We ground these results in the
established mathematics of measurement (Fisher--Rao information geometry, sloppy models, principal curves,
Backus--Gilbert resolution theory), from which the lawfulness is expected, and argue that the contribution
is the predictive, catalog-validated \emph{application} to gravitational-wave posteriors and its use as a
systematics diagnostic and an independent test of gravity.
\end{abstract}

\section{Introduction}
Gravitational-wave (GW) parameter estimation delivers, for each detected compact-binary coalescence, a
posterior probability distribution over the source parameters. It is a commonplace that the chirp mass
$\mathcal{M}_c = (m_1 m_2)^{3/5}/(m_1+m_2)^{1/5}$ is the best-measured mass combination and that the
component-mass posterior forms an elongated ``banana'' lying along near-constant-$\mathcal{M}_c$ contours
\citep{CutlerFlanagan1994,PoissonWill1995,Hannam2013}. This paper asks a sharper question: is the
\emph{geometry} of that posterior --- specifically the orientation of its principal axis --- quantitatively
\emph{predictable}, event by event, with no free parameters, and does the prediction survive out-of-sample
across independent catalogs?

We answer yes. The organizing thesis is that the uncertainty geometry of GW inference is \emph{measurement
optics}: it is a property of what the instrument and signal can and cannot resolve, largely independent of
source astrophysics. This viewpoint is not new mathematically --- it is the content of Fisher--Rao
information geometry \citep{Amari2016}, the ``sloppy model'' framework \citep{Transtrum2015}, principal
curves \citep{HastieStuetzle1989}, and Backus--Gilbert resolution theory \citep{BackusGilbert1968}. Our
contribution is to make it \emph{predictive and validated} for GW posteriors, and to turn it into two tools:
a systematics detector and an independent, waveform-fit-free test of general relativity (GR).

\section{The curved chirp-mass law}
\subsection{Definition}
For an event with source-frame samples $\{(m_1,m_2)\}$, let $\psi_{\rm meas}$ be the position angle of the
principal axis of the sample covariance, and let the axis ratio $\mathrm{axr}$ be the ratio of its
principal standard deviations. The constant-$\mathcal{M}_c$ \emph{curve} through the plane is
\begin{equation}
m_1(q) = \mathcal{M}_c\,(1+q)^{1/5} q^{-3/5}, \qquad m_2 = q\,m_1,
\end{equation}
where $q=m_2/m_1\in(0,1]$ and $\mathcal{M}_c$ is the event's median source-frame chirp mass. The predicted
orientation $\psi_{\rm curve}$ is the principal axis of this curve evaluated over the event's own $q$
marginal. The local linear approximation $\psi_{\rm tangent}$ is the tangent to the contour at the median
point. There are no calibration constants: every ingredient is supplied by the event.

\subsection{Out-of-sample validation across three catalogs}
The law was derived from GWTC-1/2.1/3 events and its prediction placed on record before any later-catalog
file was opened. Table~\ref{tab:curve} reports the locked decision on two subsequent, disjoint catalogs.
On elongated events ($\mathrm{axr}\ge 3$, where orientation is well defined) the median mismatch
$|\psi_{\rm curve}-\psi_{\rm meas}|$ is $1.26^\circ$ (GWTC-4.0/O4a) and $1.22^\circ$ (GWTC-5.0/O4b); the
curve beats the tangent on the full catalogs, and the tangent-law replication statistics reproduce the
training-set values. The catalogs are mutually disjoint (zero shared events).

\begin{table}[h]
\centering
\begin{tabular}{lccc}
\hline
 & GWTC-3 (train) & GWTC-4.0/O4a & GWTC-5.0/O4b \\
\hline
median $|\psi_{\rm curve}-\psi_{\rm meas}|$ (elongated) & --- & $\mathbf{1.26^\circ}$ & $\mathbf{1.22^\circ}$ \\
median $|\psi_{\rm curve}-\psi_{\rm meas}|$ (all) & --- & $7.53^\circ$ & $5.41^\circ$ \\
median $|\psi_{\rm tangent}-\psi_{\rm meas}|$ (all) & $7.49^\circ$ & $9.36^\circ$ & $8.23^\circ$ \\
chirp-vs-total win fraction & $81\%$ & $86\%$ & $88\%$ \\
Spearman$(|\Delta\psi_{\rm tangent}|,\mathrm{axr})$ & $-0.42$ & $-0.40$ & $-0.69$ \\
$N$ events (elongated) & --- & $86\ (19)$ & $104\ (32)$ \\
\hline
\end{tabular}
\caption{The curved chirp-mass law, locked before data and scored out-of-sample. The curve prediction
lands within $\sim1.2^\circ$ of the measured orientation on elongated events across two independent
catalogs; the tangent leaves a systematic residual that grows with elongation (negative Spearman).}
\label{tab:curve}
\end{table}

\section{Why it is lawful: the mathematics of measurement}
The law is a consequence, not a coincidence. The GW inspiral phase is dominated at leading (0PN) order by
$\mathcal{M}_c$, so the Fisher information has one sharply-measured (``stiff'') direction --- constant
$\mathcal{M}_c$ --- and poorly-measured (``sloppy'') directions transverse to it \citep{Transtrum2015}. By
the Bernstein--von Mises theorem the posterior tends to a Gaussian with covariance equal to the inverse
Fisher matrix \citep{Amari2016}, so its long axis aligns with the constant-$\mathcal{M}_c$ level set; by
{\v C}encov's theorem the Fisher--Rao metric is the unique metric invariant under sufficient statistics, and
$\mathcal{M}_c$ is approximately sufficient for the inspiral. The distinction between the tangent and the
curve is precisely the distinction between linear principal-component analysis and a \emph{principal curve}
\citep{HastieStuetzle1989}: a chord of a curved level set deviates from its tangent by an amount that grows
with the chord length, which is the observed elongation dependence. The ``resolved vs.\ unresolved
directions'' of Backus--Gilbert resolution theory \citep{BackusGilbert1968} are the same split in the
inverse-problem language. To our knowledge, none of this literature predicts the per-event $(m_1,m_2)$
posterior orientation from the chirp-mass curve, decomposes the tangent-vs-curve residual against
elongation, or validates such a law out-of-sample across catalogs; that predictive application is the
contribution here.

\section{Coherence as a systematics detector}
If the geometry is measurement optics, then \emph{departures} from it are informative, and departures that
are \emph{coherent} across events are systematics rather than physics. We test this in two ways. First, a
physics-blind ``outlier atlas'': ranking the elongated O4b events by their curve residual and correlating
against nine pre-declared axes (precession $\chi_p$, aligned spin $\chi_{\rm eff}$, mass ratio, redshift,
signal-to-noise ratio, chirp mass, total mass, secondary mass, and waveform-family disagreement) with
Benjamini--Hochberg control and a partial correlation removing the axis-ratio confound, \emph{no physical
axis} survives; the only surviving correlate is waveform-model disagreement --- a measurement systematic,
with a sign indicating that the curve tracks the waveform \emph{ensemble} consensus better than any single
waveform tracks another. Second, and more sharply, the geometric GR-exponent measurement of
Section~\ref{sec:gr} exhibits a coherent $3\sigma$ ``deviation'' that the same lens correctly identifies as
a systematic. The methodological rule --- an anomaly coherent across independent datasets is a systematic,
not a signal --- is thus both stated and exercised.

\section{A geometric test of general relativity}
\label{sec:gr}
The exponents of $\mathcal{M}_c$ are fixed by the 0PN quadrupole formula. Generalize to the one-parameter
family
\begin{equation}
\mathcal{M}_p = (m_1 m_2)^p (m_1+m_2)^{1-2p}, \qquad \text{GR: } p = 3/5 = 0.600,
\end{equation}
and \emph{measure} $p$ from the posterior orientations: because orientation is scale-free
($m_1(q;p)\propto q^{-p}(1+q)^{2p-1}$), the fit isolates the exponent from the mass scale. A deviation would
signal modified inspiral phasing (e.g.\ dipole radiation). On the O4b elongated events alone we find
$p = 0.616\pm0.011$; combining O4a and O4b, $p = 0.628\pm0.009$.

Taken naively this is a coherent $3.1\sigma$ departure from GR --- and it is a false alarm. Two
pre-committed checks demote it. (i) The two independent catalogs disagree with each other by $0.031$, as
large as the offset from GR ($0.028$); this catalog-to-catalog spread is a systematic that the
bootstrap statistical error ($0.009$) does not capture. Including it, $p = 0.628\pm0.009\,(\mathrm{stat})\pm
0.016\,(\mathrm{syst})$, or $1.5\sigma$ from GR. (ii) The per-event exponent depends on elongation
(Spearman $= -0.41$, coherent in both catalogs); the cleanest, most-elongated half yields $p=0.606$,
consistent with GR. The offset is therefore a finite-width imperfection of the leading-order geometric
model, coherent across catalogs, and vanishing in the clean limit --- not a violation of GR. We stress that
this test is a methodologically \emph{independent} cross-check, not more sensitive than direct
parameterized-deviation bounds; its value is that it shares none of the waveform-template machinery.

\section{The information anatomy of a detection}
The same Fisher structure that fixes the posterior orientation also fixes \emph{when}, in frequency, each
parameter's information arrives. Computing the cumulative Fisher information for the chirp mass (0PN), mass
ratio (1PN) and spin (1.5PN) across all $190$ real O4a+O4b events, the ordering $\mathcal{M}_c<\eta<\chi$
holds universally: the chirp mass is always determined earliest ($\sim35$~Hz), spin latest, and the
frequency separation --- the anatomy's ``richness'' --- is a monotonic function of total mass (Spearman
$-1.00$), widest for light, long-inspiral systems and compressed for heavy ones that merge immediately in
band. This is a frequency-resolved, per-event map of what each detection measured and when. It is a forward
prediction from the waveform model; a fully data-driven version (re-analysing loud events in frequency
slices and comparing empirical to GR-predicted information accumulation) is left to future work.

\section{The loudest event: GW250114}
As an application, GW250114 --- the loudest event to date (network SNR $\sim80$), a near-equal-mass binary
black hole with remnant $M_f^{\rm src}\approx62.7\,M_\odot$, $\chi_f\approx0.68$ --- is essentially a much
louder counterpart of the first detection GW150914. Its remnant predicts Kerr quasi-normal-mode frequencies
of $248.9$~Hz ($\tau=4.10$~ms) for the fundamental and $243.4$~Hz ($\tau=1.35$~ms) for the first overtone;
published ringdown spectroscopy detects the overtone at $99.9\%$ confidence and finds no deviation from the
Kerr prediction \citep{Suzuki2026}, confirming the no-hair theorem at record signal-to-noise. Being
near-equal-mass, its posterior is nearly round and it is (by construction) outside the elongated set used
for the curved law --- an illustration that the loudest events are not always the most informative for a
given geometric probe.

\section{Discussion}
We have shown that the mass-plane posterior geometry of GW events obeys a predictive, zero-parameter
constant-chirp-mass curve law, validated out-of-sample across three catalogs; that its residuals are
physics-blind (a systematics detector, not a source-class probe); and that inverting it yields an
independent geometric test of GR's inspiral mass structure, consistent with general relativity once
systematics are controlled. The results rest on well-understood measurement mathematics, which is a
strength: the lawfulness is expected, and the value lies in the predictive application, the out-of-sample
discipline, and the diagnostic uses.

\paragraph{Honest limitations.} (1) The curved law is validated on elongated events; round posteriors have
ill-defined orientation and are excluded by design. (2) The geometric GR test is a cross-check, less
sensitive than dedicated waveform-based bounds, and is currently systematics-limited by a finite-width
model imperfection. (3) The information anatomy is model-predicted, not strain-extracted. (4) The
first-principles derivation of the curve's exact width and curvature from the Fisher information remains
open. (5) Waveform-model choice contributes a per-event orientation scatter of a few degrees; it is
incoherent and averages out of population-level quantities, but is the dominant per-event systematic.

\paragraph{Data availability.} All data are public: GWTC-4.0 PE (Zenodo 16053484) and GWTC-5.0 PE (Zenodo
20276106, 20348006). Analysis code, locked pre-registrations, per-event results, and a reproducibility
notebook accompany this work.

\begin{thebibliography}{99}
\bibitem[Cutler \& Flanagan(1994)]{CutlerFlanagan1994} C.~Cutler and {\'E}.~{\'E}.~Flanagan,
\emph{Phys.\ Rev.\ D} \textbf{49}, 2658 (1994).
\bibitem[Poisson \& Will(1995)]{PoissonWill1995} E.~Poisson and C.~M.~Will,
\emph{Phys.\ Rev.\ D} \textbf{52}, 848 (1995).
\bibitem[Baird et al.(2013)]{Baird2013} E.~Baird, S.~Fairhurst, M.~Hannam, and P.~Murphy,
\emph{Phys.\ Rev.\ D} \textbf{87}, 024035 (2013).
\bibitem[Hannam et al.(2013)]{Hannam2013} M.~Hannam, D.~A.~Brown, S.~Fairhurst, C.~L.~Fryer, and
I.~W.~Harry, \emph{Astrophys.\ J.\ Lett.} \textbf{766}, L14 (2013).
\bibitem[Ohme et al.(2013)]{Ohme2013} F.~Ohme, A.~B.~Nielsen, D.~Keppel, and A.~Lundgren,
\emph{Phys.\ Rev.\ D} \textbf{88}, 042002 (2013).
\bibitem[Fairhurst et al.(2023)]{Fairhurst2023} S.~Fairhurst, C.~Hoy, R.~Green, C.~Mills, and
S.~A.~Usman, \emph{Phys.\ Rev.\ D} \textbf{108}, 082006 (2023).
\bibitem[Amari(2016)]{Amari2016} S.~Amari, \emph{Information Geometry and Its Applications}
(Springer, 2016).
\bibitem[Transtrum et al.(2015)]{Transtrum2015} M.~K.~Transtrum et al.,
\emph{J.\ Chem.\ Phys.} \textbf{143}, 010901 (2015).
\bibitem[Hastie \& Stuetzle(1989)]{HastieStuetzle1989} T.~Hastie and W.~Stuetzle,
\emph{J.\ Amer.\ Statist.\ Assoc.} \textbf{84}, 502 (1989).
\bibitem[Backus \& Gilbert(1968)]{BackusGilbert1968} G.~Backus and F.~Gilbert,
\emph{Geophys.\ J.\ R.\ Astron.\ Soc.} \textbf{16}, 169 (1968).
\bibitem[Suzuki et al.(2026)]{Suzuki2026} M.~Suzuki, K.~Kubota, S.~Morisaki, H.~Motohashi, and
D.~Watarai, arXiv:2605.03576 (2026).
\end{thebibliography}

\end{document}
